---
title: "PGP Assay Analysis"
author: "Shawn Higdon"
date: "11/8/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries
```{r, message=FALSE}
library(treeio)
library(ggtree)
library(ggnewscale)
library(ComplexHeatmap)
library(tidyverse)
library(dendextend)
library(ape)
library(circlize)
library(viridis)
library(randomcoloR)
library(RColorBrewer)
library(cowplot)
```

# Read in Genome Metadata

> ABB to BCW Genome ID Map

> Metabat Bin Count per Isolate

> Sourmash LCA Classification Data

> Genome Assembly Statistics: Quast output and coverage estimation

```{r}
# ABB to BCW ID MAP
abb_bcw_labels <- read.csv("./genome_metadata/abb_genome_bcw_labels.csv", header = T)

# METABAT BIN INFO
metabat_bins <- read_csv("./genome_metadata/metabat_bin_count.csv", col_names = c("ID", "n_bins"))

# Sourmash LCA Genome Classification Data
smash_lca_whole_genomes <- read.csv("./sourmash_data/sm-lca-whole_genomes-all-k31.csv", header = T)

# Read in quast Summary for all isolate assemblies done with MEgahit
asm_stats <- read_tsv("genome_metadata/assembly_stats/all_quast_reports.tsv") # includes eukaryotic isolates

# Read in assembly coverage stat table
asm_cov <- read.table("genome_metadata/assembly_stats/avg_coverage.tsv", header = T)

# Add AVG_COV to asm_stats
asm_stats <- inner_join(asm_stats, asm_cov, by = "Assembly")

```

## Add BCW labels to all metadata files
```{r}
# Add BCW label to metabat bin DF
metabat_bins$bcw_id <- abb_bcw_labels$BCW_ID[match(metabat_bins$ID, abb_bcw_labels$ABB_ID)]

# Add BCW label to sourmash lca data
smash_lca_whole_genomes$bcw_id <- abb_bcw_labels$BCW_ID[match(smash_lca_whole_genomes$ID, abb_bcw_labels$ABB_ID)]

## change '_' to '-' for bcw_id
smash_lca_whole_genomes$bcw_id <- sub("_", "-", smash_lca_whole_genomes$bcw_id)

## empty cells are 'unassigned'
smash_lca_whole_genomes$genus <- sub("^$", "unassigned", smash_lca_whole_genomes$genus)

```

## Generate list of bacterial isolates with single genome bin
```{r}
# create new dataframe
single_isolates <- metabat_bins %>% select( "bcw_id", "ID", "n_bins")
## modify to match other records
single_isolates$bcw_id <- sub("_", "-", single_isolates$bcw_id)

# add genus data
single_isolates$genus <- smash_lca_whole_genomes$genus[match(single_isolates$bcw_id, smash_lca_whole_genomes$bcw_id)]

# add phylum data
single_isolates$phylum <- smash_lca_whole_genomes$phylum[match(single_isolates$bcw_id, smash_lca_whole_genomes$bcw_id)]

# empty cells for genus are 'unassigned'
single_isolates$genus <- sub("^$", "unassigned", single_isolates$genus)

# empty cells for Phylum are 'unassigned'
single_isolates$phylum <- sub("^$", "unassigned", single_isolates$phylum)

# remove eukaryotes
single_isolates <- single_isolates %>% filter(genus != "Meyerozyma" &
                                                genus != "Rhodotorula")
# remove isolates that were pulled apart
single_isolates <- single_isolates %>% filter(bcw_id != "BCW_201831.1" &
                                                bcw_id != "BCW_201831.2" &
                                                bcw_id != "BCW_200557.1" &
                                                bcw_id != "BCW_200557.2")
# Keep only isolates with 1 genomic bin
single_isolates <- single_isolates %>% filter(n_bins == 1)

# add abb_ids
single_isolates$abb_id <- abb_bcw_labels$ABB_ID[match(single_isolates$bcw_id, abb_bcw_labels$BCW_ID)]

# change '_' to '-' for bcw_id
single_isolates$bcw_id <- sub("_", "-", single_isolates$bcw_id)

```

## PGP Analysis Data

### Read in the Data for PGP Biochemical assays

```{r, message=FALSE}

# Read in the PGP Assay raw data files for each isolate
pgp_data_list <- list.files(path = "./PGP_Data", pattern = 'PGP_*', recursive = T, full.names = T)
pgp_data_list

# Create a list of bcw nifscan search output tables
pgp_data_tbl_list <- lapply(pgp_data_list, read_csv)

# create object for each PGP Assay
pgp_15N <- pgp_data_tbl_list[[1]]
pgp_accd <- pgp_data_tbl_list[[2]]
pgp_auxin <- pgp_data_tbl_list[[3]]
pgp_po4 <- pgp_data_tbl_list[[4]]
pgp_sidero <- pgp_data_tbl_list[[5]]

## change '_' to '-' for bcw_id of pgp_15N
pgp_15N$BCW_ID <- sub("_", "-", pgp_15N$BCW_ID)
```


## Compute PGP Assay metrics

### ACC Deaminase Growth Assay

> Isolates cultured on medium containing 1-amino-1-cyclopropane carboxylic acid as the only source of available nitrogen

```{r}
# ACCD Assay
## Compute average for Day 0 OD600 replicates
pgp_accd$OD600_D0_avg <- rowMeans(pgp_accd[,3:5], na.rm = TRUE)
## Compute average for Day 4 OD600 replicates
pgp_accd$OD600_D4_avg <- rowMeans(pgp_accd[,6:8], na.rm = TRUE)

## Compute Relative growth rate for each isolate
pgp_accd$ACCD_RGR <- pgp_accd$OD600_D4_avg/pgp_accd$OD600_D0_avg

## change number of significant digits for numeric values to 3
pgp_accd %>% mutate_if(is.numeric, round, digits = 3)

## change '_' to '-' for bcw_id
pgp_accd$BCW_ID <- sub("_", "-", pgp_accd$BCW_ID)

## inspect object
str(pgp_accd)
```

### Auxin Biosynthesis Assay

```{r}
# IAA Assay
## Compute average for OD535 replicates
pgp_auxin$OD535_avg <- rowMeans(pgp_auxin[,3:5], na.rm = TRUE)

## Compute average for auxin concentration replicate values
pgp_auxin$IAA_mg_ml_avg <- rowMeans(pgp_auxin[,6:8], na.rm = TRUE)

## change number of significant digits for numeric values to 3
pgp_auxin %>% mutate_if(is.numeric, round, digits = 3)

## change '_' to '-' for bcw_id
pgp_auxin$BCW_ID <- sub("_", "-", pgp_auxin$BCW_ID)

## Inspect object
str(pgp_auxin)
```

## Phosphate Solubilization Assay

```{r}
# PO4 Assay
## Compute average for OD535 replicates
pgp_po4$OD535_avg <- rowMeans(pgp_po4[,3:5], na.rm = TRUE)

## Compute average for auxin concentration replicate values
pgp_po4$PO4_mg_L_avg <- rowMeans(pgp_po4[,6:8], na.rm = TRUE)

## Multiply diluted concentrations by assay dilution factor of 100
pgp_po4$PO4_mg_L_avg <- pgp_po4$PO4_mg_L_avg*100

## change number of significant digits for numeric values to 3
pgp_po4 %>% mutate_if(is.numeric, round, digits = 3)

## change '_' to '-' for bcw_id
pgp_po4$BCW_ID <- sub("_", "-", pgp_po4$BCW_ID)

## Inspect object
str(pgp_po4)
```

## Siderophore Production Assay

```{r}
# Siderophore Assay
## Compute average for OD535 replicates
pgp_sidero$OD630_avg <- rowMeans(pgp_sidero[,3:5], na.rm = TRUE)

## Compute average for Siderophore units (%) using AR_AS replicate values
## % Siderophore units = Absorbance of Sample (AS) - Absorbance of Reference (AR) / AR
pgp_sidero$Sidero_units_avg <- rowMeans(pgp_sidero[,6:8], na.rm = TRUE)

## change number of significant digits for numeric values to 3
pgp_sidero %>% mutate_if(is.numeric, round, digits = 3)

## change '_' to '-' for bcw_id
pgp_sidero$BCW_ID <- sub("_", "-", pgp_sidero$BCW_ID)

## Inspect Object
str(pgp_sidero)
```

### Sourmash Compare Data
```{r}
# read in sourmash k31 compare matrix for single isolate genomes
sm_k31_cmp <- read.csv("sourmash_data/pure_isolate_492_cmp/pure_isolate_k31_cmp_492.csv", header=T, check.names = F)

# set rownames to colnames
rownames(sm_k31_cmp) <- colnames(sm_k31_cmp)
#head(rownames(sm_k31_cmp))

# add bcw ids to sm_k31_cmp data frame

## change '_' to '-' for BCW ids
abb_bcw_labels$BCW_ID <- sub("_", "-", abb_bcw_labels$BCW_ID)

## match BCW ids
sm_k31_cmp$bcw_id <- abb_bcw_labels$BCW_ID[match(rownames(sm_k31_cmp), abb_bcw_labels$BCW_ID)]

# add genus info to sm_k31
sm_k31_cmp$genus <- single_isolates$genus[match(sm_k31_cmp$bcw_id, single_isolates$bcw_id)]

# Create tx_name Variable with Hybrid genus + Isolate Name: GENUS sp. BCWXXXX
sm_k31_cmp$tx_name <- paste(sm_k31_cmp$genus, sprintf("sp. %s", sm_k31_cmp$bcw_id))

# create matrix
sm_k31_cmp_mat <- as.matrix(sm_k31_cmp[,1:492])

rownames(sm_k31_cmp_mat) <- sm_k31_cmp$tx_name
colnames(sm_k31_cmp_mat) <- sm_k31_cmp$tx_name

```

## Combine PGP Assay Results

> Make a data frame with results from:

1. 15N Assay

2. ACCD Growth Assay

3. Auxin Biosynthesis Assay

4. Phosphate Solubilization Assay

5. Siderophore Production Assay

### PGP Master Data Frame
```{r}
# Make pgp master dataframe
pgp_master <- data.frame(TX_ID=sm_k31_cmp$tx_name)

# add BCW ids
pgp_master$BCW_ID <- sm_k31_cmp$bcw_id

# add ABB ids
pgp_master$ABB_ID <- abb_bcw_labels$ABB_ID[match(pgp_master$BCW_ID, abb_bcw_labels$BCW_ID)]

# add N15 ratio values for each isolate
pgp_master$`BNF` <- pgp_15N$ratio[match(pgp_master$BCW_ID, pgp_15N$BCW_ID)]

# add ACCD assay results
pgp_master$`ACC` <- pgp_accd$ACCD_RGR[match(pgp_master$BCW_ID, pgp_accd$BCW_ID)]

# add Auxin assay results
pgp_master$`IAA` <- pgp_auxin$IAA_mg_ml_avg[match(pgp_master$BCW_ID, pgp_auxin$BCW_ID)]

# add Phosphate solubility assay results
pgp_master$`PO4` <- pgp_po4$PO4_mg_L_avg[match(pgp_master$BCW_ID, pgp_po4$BCW_ID)]

# add siderophore assay results
pgp_master$`SID` <- pgp_sidero$Sidero_units_avg[match(pgp_master$BCW_ID, pgp_sidero$BCW_ID)]

# add Genus classification for each isolate
pgp_master$Genus <- smash_lca_whole_genomes$genus[match(pgp_master$BCW_ID, smash_lca_whole_genomes$bcw_id)]

pgp_master$Genus <- as.factor(pgp_master$Genus) # coerce to factor

## assign a unique color to each genus
set.seed(8888)
pgp_genus_colors <- data.frame(Genus = unique(pgp_master$Genus), Color =  distinctColorPalette(length(unique(pgp_master$Genus))))

# add phylum classification for each isolate
pgp_master$Phylum <- single_isolates$phylum[match(pgp_master$BCW_ID, single_isolates$bcw_id)]

## assign a unique color to each phylum
set.seed(1234)
pgp_phylum_colors <- data.frame(Phylum = unique(pgp_master$Phylum), Color =  distinctColorPalette(length(unique(pgp_master$Phylum))))

### match Genus
pgp_master$Genus_Color <- pgp_genus_colors$Color[match(pgp_master$Genus, pgp_genus_colors$Genus)]

### match Phylum
pgp_master$Phylum_Color <- pgp_phylum_colors$Color[match(pgp_master$Phylum, pgp_phylum_colors$Phylum)]

## change number of significant digits for numeric values to 3
pgp_master <- pgp_master %>% mutate_if(is.numeric, round, digits = 3)

# Remove incomplete cases (Isolates with NA values for PGP assays)
#pgp_master <- drop_na(pgp_master)

# Create list of isolates with complete pgp cases
single_isolates_complete_cases <- data.frame(BCW_ID = pgp_master$BCW_ID)
#write_csv(single_isolates_complete_cases, "./R_output_files/single_isolates_complete_cases.csv", col_names = F)

# add sm_k31_cmp data to pgp_master
pgp_master <- cbind(sm_k31_cmp_mat, pgp_master)

```

> 492 isolates with single bin genomes and complete cases for pgp assays.

> Used list of single genome bin with complete cases to perform `sourmash compare` with subset of isolate genomes.

> read in output from `sourmash compare` to R

```
sourmash compare -k 31
```
### All Single Isolates
```{r}
# Create Phylo Tree from sourmash matrix

sm_k31_tree_mat <- dist2(sm_k31_cmp_mat)

sm_k31_tree_fit <- hclust(sm_k31_tree_mat)

sm_k31_phylo <- as.phylo(sm_k31_tree_fit)

# set rownames of pgp_master to match tree tip labels
rownames(pgp_master) <- sm_k31_phylo$tip.label

head(rownames(pgp_master)) # inspect rownames

colnames(pgp_master[,493:504]) # show column names

levels(pgp_master$Genus) # assess levels of genera

```

#### PGP Phenotype Tree
```{r, message=FALSE}
# Make tree
sm_k31_tree_o15 <- ggtree(sm_k31_phylo, color = "black", size = 0.3, layout = "fan", open.angle = 15)

# Genus Tree
pgp0_notip <- gheatmap(sm_k31_tree_o15, pgp_master[, "Genus", drop = F],
                     offset = 0,
                     width = 0.05,
                     colnames_position = "top",
                     colnames_angle = 0,
                     colnames_offset_y = 9.5,
                     font.size = 4,
                     color = "grey50") + 
  scale_fill_manual(values=c("#E6BE53",
                             "#CAD773",
                             "#81D870",
                             "#8F3EE4",
                             "#E25F9D",
                             "#D8E9D6",
                             "#72A890",
                             "#CDEC4C",
                             "#C59D6B",
                             "#86D6E1",
                             "#77E94F",
                             "#81748E",
                             "#64E7A9",
                             "#D3D0E3",
                             "#767CDE",
                             "#7DAEDE",
                             "#E2684F",
                             "#E1ABAA",
                             "#7BEFDF",
                             "#DA47D2",
                             "#D7A7DF",
                             "#E5DEA8",
                             "#B4E3A7",
                             "#CF7DDC"),
                    name = "Genus",
                    guide = guide_legend(
                      direction = "vertical",
                      ncol = 2,
                      title.position = "top",
                      title.theme = element_text(size = 18, hjust = 0.5),
                      label.position = "right",
                      label.theme = element_text(size = 14)
                      )) + theme(legend.title = element_text(),
                                 legend.key.size = unit(.5, "cm"))

pgp_p0 <- pgp0_notip + new_scale_fill()

# Add gheatmap 1 for ACCD Utilization Data

pgp_p1 <- gheatmap(pgp_p0, pgp_master[, "ACC", drop = F],
                     offset = 0.25,
                     width = 0.05,
                     colnames_position = "top",
                     colnames_angle = 0,
                     colnames_offset_y = 9.5,
                     font.size = 4,
                     color = "grey50") + 
     scale_fill_viridis_c(option = "viridis",
                       direction = -1,
                       name = "ACC (Relative Growth Rate)",
                       limits = c(0, 5),
                       breaks = c(0, 1, 2, 3, 4, 5),
                       guide = guide_colorbar(
                         direction = "horizontal",
                         title.position = "top",
                         title.theme = element_text(size = 18, hjust = 0.5),
                         barwidth = 18,
                         label.position = "bottom",
                         label.theme = element_text(size = 14)
                         )
                       ) +
  theme(legend.title = element_text(), legend.position = "right")

# Add gheatmap 2 for BNF assay data
pgp_p2 <- pgp_p1 + new_scale_fill()

pgp_p3 <- gheatmap(pgp_p2, pgp_master[, "BNF", drop=F],
                 offset = 0.45,
                 width=0.05,
                 colnames_angle = 0,
                 colnames_position = "top",
                 colnames_offset_y = 9.5,
                 font.size = 4,
                 color = "grey50") + 
  scale_fill_viridis_c(option = "inferno",
                       direction = -1,
                       name = "BNF (15N/14N)",
                       limits = c(0, 5),
                       breaks = c(0, 1, 2, 3, 4, 5),
                       guide = guide_colorbar(
                         direction = "horizontal",
                         title.position = "top",
                         title.theme = element_text(size = 18, hjust = 0.5),
                         barwidth = 18,
                         label.position = "bottom",
                         label.theme = element_text(size = 14)
                         )
                       ) +
  theme(legend.title = element_text(), legend.position = "right")

# Add gheatmap 3 for IAA Biosynthesis assay
pgp_p4 <- pgp_p3 + new_scale_fill()

pgp_p5 <- gheatmap(pgp_p4, pgp_master[, "IAA", drop=F],
                    offset = 0.65,
                    width = 0.05,
                    colnames_angle = 0,
                    colnames_position = "top",
                    colnames_offset_y = 9.5,
                    font.size = 4,
                    color = "grey50") + 
  scale_fill_viridis_c(option = "plasma",
                       direction = 1,
                       name = "IAA (mg/mL)",
                       limits = c(-5, 215),
                       breaks = c(0, 50, 100, 150, 200),
                       guide = guide_colorbar(
                         direction = "horizontal",
                         title.position = "top",
                         title.theme = element_text(size = 18, hjust = 0.5),
                         barwidth = 18,
                         label.position = "bottom",
                         label.theme = element_text(size = 14)
                         )
                       ) +
  theme(legend.title = element_text(), legend.position = "right")

# Add gheatmap 3 for assay 3
pgp_p6 <- pgp_p5 + new_scale_fill()

pgp_all <- gheatmap(pgp_p6, pgp_master[, "PO4", drop=F],
                    offset = 0.85,
                    width = 0.05,
                    colnames_angle = 0,
                    colnames_position = "top",
                    colnames_offset_y = 9.5,
                    font.size = 4,
                    color = "grey50") + 
  scale_fill_viridis_c(option = "magma",
                       direction = 1,
                       name = "PO4 (mg/L)",
                       limits = c(0, 2250),
                       breaks = c(0, 500, 1000, 1500, 2000),
                       guide = guide_colorbar(
                         direction = "horizontal",
                         title.position = "top",
                         title.theme = element_text(size = 18, hjust = 0.5),
                         barwidth = 18,
                         label.position = "bottom",
                         label.theme = element_text(size = 14)
                         )
                       ) +
  theme(legend.title = element_text(), legend.position = "right")

pgp_all_rot <- print(rotate_tree(pgp_all, -82))

# save plot
ggsave("./Plots/pgp_all_rot.pdf", pgp_all_rot, width = 22, height = 14)

```

### NIF Groups

> Read Sourmash K31 Compare Data for Each NIF Group

### Dos-Santos Positive (DSP) - Group A
```{r}
# DSP Group
smash_dsp_cmp <- read.csv("./sourmash_data/dsp_dos-santos-positive/pure_iso_dsp_k31_cmp.csv", header = TRUE, check.names = FALSE)

# set rownames to colnames
rownames(smash_dsp_cmp) <- colnames(smash_dsp_cmp)

# add bcw ids to smash_dsp_cmp data frame
smash_dsp_cmp$bcw_id <- abb_bcw_labels$BCW_ID[match(rownames(smash_dsp_cmp), abb_bcw_labels$BCW_ID)]

# add genus info to smash_dsp_cmp
smash_dsp_cmp$Genus <- smash_lca_whole_genomes$genus[match(smash_dsp_cmp$bcw_id, smash_lca_whole_genomes$bcw_id)]

# add tx_name Variable to smash_dsp_cmp
smash_dsp_cmp$tx_name <- paste(smash_dsp_cmp$Genus, sprintf("sp. %s", smash_dsp_cmp$bcw_id))

# add BNF value
smash_dsp_cmp$BNF <- pgp_15N$ratio[match(smash_dsp_cmp$bcw_id, pgp_15N$BCW_ID)]

# add ACC value
smash_dsp_cmp$ACC <- pgp_accd$ACCD_RGR[match(smash_dsp_cmp$bcw_id, pgp_accd$BCW_ID)]

# add IAA value
smash_dsp_cmp$IAA <- pgp_auxin$IAA_mg_ml_avg[match(smash_dsp_cmp$bcw_id, pgp_auxin$BCW_ID)]

# add PO4 value
smash_dsp_cmp$PO4 <- pgp_po4$PO4_mg_L_avg[match(smash_dsp_cmp$bcw_id, pgp_po4$BCW_ID)]

# set rownames to match sample
rownames(smash_dsp_cmp) <- smash_dsp_cmp$tx_name

# create matrices for tree

## sourmash compare matrix
smash_dsp_cmp_mat <- as.matrix(smash_dsp_cmp[,1:193]) # 193 isolates
rownames(smash_dsp_cmp_mat) <- smash_dsp_cmp$tx_name
colnames(smash_dsp_cmp_mat) <- rownames(smash_dsp_cmp_mat)

```

#### DSP tree objects
```{r}
# Create Phylo Tree from sourmash matrix

smash_dsp_k31_tree_mat <- dist2(smash_dsp_cmp_mat)
smash_dsp_k31_tree_fit <- hclust(smash_dsp_k31_tree_mat)
smash_dsp_k31_phylo <- as.phylo(smash_dsp_k31_tree_fit)

```


#### DSP ggtree
```{r}
# Make tree
sm_dsp_k31_tree <- ggtree(smash_dsp_k31_phylo,
                          color = "black",
                          size = 0.4,
                          layout = "fan",
                          open.angle = 15)
```

### Semi Dos-Santos (SDS) - Group B
```{r}
# SDS Group
smash_sds_cmp <- read.csv("./sourmash_data/sds_semi-dos-santos/pure_iso_sds_k31_cmp.csv", header = TRUE, check.names = FALSE)

# set rownames to colnames
rownames(smash_sds_cmp) <- colnames(smash_sds_cmp)

# add bcw ids to smash_sds_cmp data frame
smash_sds_cmp$bcw_id <- abb_bcw_labels$BCW_ID[match(rownames(smash_sds_cmp), abb_bcw_labels$BCW_ID)]

# add genus info to smash_sds_cmp
smash_sds_cmp$Genus <- single_isolates$genus[match(smash_sds_cmp$bcw_id, single_isolates$bcw_id)]

# add tx_name Variable to smash_sds_cmp
smash_sds_cmp$tx_name <- paste(smash_sds_cmp$Genus, sprintf("sp. %s", smash_sds_cmp$bcw_id))

# add BNF value
smash_sds_cmp$BNF <- pgp_15N$ratio[match(smash_sds_cmp$bcw_id, pgp_15N$BCW_ID)]

# add ACC value
smash_sds_cmp$ACC <- pgp_accd$ACCD_RGR[match(smash_sds_cmp$bcw_id, pgp_accd$BCW_ID)]

# add IAA value
smash_sds_cmp$IAA <- pgp_auxin$IAA_mg_ml_avg[match(smash_sds_cmp$bcw_id, pgp_auxin$BCW_ID)]

# add PO4 value
smash_sds_cmp$PO4 <- pgp_po4$PO4_mg_L_avg[match(smash_sds_cmp$bcw_id, pgp_po4$BCW_ID)]

# set rownames to match sample
rownames(smash_sds_cmp) <- smash_sds_cmp$tx_name

# create matrices for tree

## sourmash compare matrix
smash_sds_cmp_mat <- as.matrix(smash_sds_cmp[,1:66]) # 66 isolates
rownames(smash_sds_cmp_mat) <- smash_sds_cmp$tx_name
colnames(smash_sds_cmp_mat) <- rownames(smash_sds_cmp_mat)

```

#### SDS tree objects
```{r}
# Create Phylo Tree from sourmash matrix

smash_sds_k31_tree_mat <- dist2(smash_sds_cmp_mat)
smash_sds_k31_tree_fit <- hclust(smash_sds_k31_tree_mat)
smash_sds_k31_phylo <- as.phylo(smash_sds_k31_tree_fit)

```


#### SDS ggtree
```{r}
# Make tree
sm_sds_k31_tree <- ggtree(smash_sds_k31_phylo,
                          color = "black",
                          size = 0.4,
                          layout = "fan",
                          open.angle = 15)
```

### Dos-Santos Negative (DSN) - Group C
```{r}
# DSN Group
smash_dsn_cmp <- read.csv("./sourmash_data/dsn_dos-santos-negative/pure_iso_dsn_k31_cmp.csv", header = TRUE, check.names = FALSE)

# set rownames to colnames
rownames(smash_dsn_cmp) <- colnames(smash_dsn_cmp)

# add bcw ids to smash_dsn_cmp data frame
smash_dsn_cmp$bcw_id <- abb_bcw_labels$BCW_ID[match(rownames(smash_dsn_cmp), abb_bcw_labels$BCW_ID)]

# add genus info to smash_dsn_cmp
smash_dsn_cmp$Genus <- single_isolates$genus[match(smash_dsn_cmp$bcw_id, single_isolates$bcw_id)]

# add tx_name Variable to smash_dsn_cmp
smash_dsn_cmp$tx_name <- paste(smash_dsn_cmp$Genus, sprintf("sp. %s", smash_dsn_cmp$bcw_id))

# add BNF value
smash_dsn_cmp$BNF <- pgp_15N$ratio[match(smash_dsn_cmp$bcw_id, pgp_15N$BCW_ID)]

# add ACC value
smash_dsn_cmp$ACC <- pgp_accd$ACCD_RGR[match(smash_dsn_cmp$bcw_id, pgp_accd$BCW_ID)]

# add IAA value
smash_dsn_cmp$IAA <- pgp_auxin$IAA_mg_ml_avg[match(smash_dsn_cmp$bcw_id, pgp_auxin$BCW_ID)]

# add PO4 value
smash_dsn_cmp$PO4 <- pgp_po4$PO4_mg_L_avg[match(smash_dsn_cmp$bcw_id, pgp_po4$BCW_ID)]

# set rownames to match sample
rownames(smash_dsn_cmp) <- smash_dsn_cmp$tx_name

# create matrices for tree

## sourmash compare matrix
smash_dsn_cmp_mat <- as.matrix(smash_dsn_cmp[,1:233]) # 233 isolates
rownames(smash_dsn_cmp_mat) <- smash_dsn_cmp$tx_name
colnames(smash_dsn_cmp_mat) <- rownames(smash_dsn_cmp_mat)

```

#### DSN tree objects
```{r}
# Create Phylo Tree from sourmash matrix

smash_dsn_k31_tree_mat <- dist2(smash_dsn_cmp_mat)
smash_dsn_k31_tree_fit <- hclust(smash_dsn_k31_tree_mat)
smash_dsn_k31_phylo <- as.phylo(smash_dsn_k31_tree_fit)

```


#### DSN ggtree
```{r}
# Make tree
sm_dsn_k31_tree <- ggtree(smash_dsn_k31_phylo,
                          color = "black",
                          size = 0.4,
                          layout = "fan",
                          open.angle = 15)

```

### Cowplot NIF Group ggtrees
##### DSP Isolates
```{r, message=FALSE}
# Add gheatmap 0 for genus
## reproducible colors
pgp0_notip_dsp <- gheatmap(sm_dsp_k31_tree, smash_dsp_cmp[, "Genus", drop = F],
                     offset = 0,
                     width = 0.05,
                     colnames_position = "top",
                     colnames_angle = 0,
                     colnames_offset_y = 3.5,
                     font.size = 3,
                     color = "grey50") + 
  scale_fill_manual(values=c("#72A890",
                             "#86D6E1",
                             "#77E94F",
                             "#E1ABAA",
                             "#7BEFDF",
                             "#DA47D2",
                             "#CF7DDC"),
                    name = "Genus",
                    guide = guide_legend(
                      direction = "vertical",
                      ncol = 2,
                      title.position = "top",
                      title.theme = element_text(size = 18, hjust = 0.5),
                      label.position = "right",
                      label.theme = element_text(size = 14)
                      )
  ) +
  theme(legend.title = element_text(), legend.key.size = unit(.5, "cm"))

# Add gheatmap 1 for ACC
pgp1_notip_dsp <- pgp0_notip_dsp + new_scale_fill()

pgp2_notip_dsp <- gheatmap(pgp1_notip_dsp, smash_dsp_cmp[, "ACC", drop = F],
                     offset = 0.2,
                     width = 0.05,
                     colnames_position = "top",
                     colnames_angle = 0,
                     colnames_offset_y = 3.5,
                     font.size = 3,
                     color = "grey50") + 
  scale_fill_viridis_c(option="viridis",
                       direction = -1,
                       name = "ACC (Relative Growth Rate)",
                       limits = c(0,5),
                       breaks = c(0, 1, 2, 3, 4, 5),
                       guide = guide_colorbar(
                         direction = "horizontal",
                         title.position = "top",
                         title.theme = element_text(size = 18, hjust = 0.5),
                         barwidth = 18,
                         label.position = "bottom",
                         label.theme = element_text(size = 14)
                         )) +
  theme(legend.title = element_text(), legend.position = "bottom")

# Add gheatmap 2 for BNF

pgp3_notip_dsp <- pgp2_notip_dsp + new_scale_fill()

pgp4_notip_dsp <- gheatmap(pgp3_notip_dsp, smash_dsp_cmp[, "BNF", drop=F],
                 offset = 0.35,
                 width=0.05,
                 colnames_angle = 0,
                 colnames_position = "top",
                 colnames_offset_y = 3.5,
                 font.size = 3,
                 color = "grey50") + 
  scale_fill_viridis_c(option="inferno",
                       direction = -1,
                       name = "BNF (15N/14N)",
                       limits = c(0,5),
                       breaks = c(0, 1, 2, 3, 4, 5),
                       guide = guide_colorbar(
                         direction = "horizontal",
                         title.position = "top",
                         title.theme = element_text(size = 18, hjust = 0.5),
                         barwidth = 18,
                         label.position = "bottom",
                         label.theme = element_text(size = 14)
                       )) +
  theme(legend.title = element_text(), legend.position = "bottom")

# Add gheatmap 3 for IAA
pgp5_notip_dsp <- pgp4_notip_dsp + new_scale_fill()

pgp6_notip_dsp <- gheatmap(pgp5_notip_dsp, smash_dsp_cmp[, "IAA", drop=F],
                    offset = 0.5,
                    width = 0.05,
                    colnames_angle = 0,
                    colnames_position = "top",
                    colnames_offset_y = 3.5,
                    font.size = 3,
                    color = "grey50") + 
  scale_fill_viridis_c(option = "plasma",
                       direction = 1,
                       name = "IAA Biosynthesis (mg/mL)",
                       limits = c(-5,215),
                       breaks = c(0, 50, 100, 150, 200),
                       guide = guide_colorbar(
                         direction = "horizontal",
                         title.position = "top",
                         title.theme = element_text(size = 18, hjust = 0.5),
                         barwidth = 18,
                         label.position = "bottom",
                         label.theme = element_text(size = 14)
                       )) +
  theme(legend.title = element_text(), legend.position = "bottom")

# Add gheatmap 4 for PO4
pgp7_notip_dsp <- pgp6_notip_dsp + new_scale_fill()

pgp_dsp_notip <- gheatmap(pgp7_notip_dsp, smash_dsp_cmp[, "PO4", drop=F],
                    offset = 0.65,
                    width = 0.05,
                    colnames_angle = 0,
                    colnames_position = "top",
                    colnames_offset_y = 3.5,
                    font.size = 3,
                    color = "grey50") + 
  scale_fill_viridis_c(option = "magma",
                       direction = 1,
                       name = "PO4 Solubilization (mg/L)",
                       limits = c(0, 2250),
                       breaks = c(0, 500, 1000, 1500, 2000),
                       guide = guide_colorbar(
                         direction = "horizontal",
                         title.position = "top",
                         title.theme = element_text(size = 18, hjust = 0.5),
                         barwidth = 18,
                         label.position = "bottom",
                         label.theme = element_text(size = 14)
                       )) +
  theme(legend.title = element_text(),
        legend.position = "bottom")

pgp_dsp_rot <- print(rotate_tree(pgp_dsp_notip, -83))

# save plot
ggsave("./Plots/pgp_dsp_rot.pdf", pgp_dsp_rot, width = 22, height = 14)

```

##### SDS Isolates
```{r, message=FALSE}
# Add gheatmap 0 for genus
## reproducible colors
pgp0_notip_sds <- gheatmap(sm_sds_k31_tree, smash_sds_cmp[, "Genus", drop = F],
                     offset = 0,
                     width = 0.05,
                     colnames_position = "top",
                     colnames_angle = 0,
                     colnames_offset_y = 0.9,
                     font.size = 3,
                     color = "grey50") + 
  scale_fill_manual(values=c("#E6BE53",
                             "#CAD773",
                             "#8F3EE4",
                             "#D8E9D6",
                             "#C59D6B",
                             "#64E7A9",
                             "#767CDE",
                             "#E1ABAA",
                             "#B4E3A7",
                             "#CF7DDC"),
                    name = "Genus",
                    guide = guide_legend(
                      direction = "vertical",
                      ncol = 2,
                      title.position = "top",
                      title.theme = element_text(size = 18, hjust = 0.5),
                      label.position = "right",
                      label.theme = element_text(size = 14)
                      )
  ) +
  theme(legend.title = element_text(), legend.key.size = unit(.5, "cm"))

# Add gheatmap 1 for ACC
pgp1_notip_sds <- pgp0_notip_sds + new_scale_fill()

pgp2_notip_sds <- gheatmap(pgp1_notip_sds, smash_sds_cmp[, "ACC", drop = F],
                     offset = 0.125,
                     width = 0.05,
                     colnames_position = "top",
                     colnames_angle = 0,
                     colnames_offset_y = 0.9,
                     font.size = 3,
                     color = "grey50") + 
  scale_fill_viridis_c(option="viridis",
                       direction = -1,
                       name = "ACC (Relative Growth Rate)",
                       limits = c(0,5),
                       breaks = c(0, 1, 2, 3, 4, 5),
                       guide = guide_colorbar(
                         direction = "horizontal",
                         title.position = "top",
                         title.theme = element_text(size = 18, hjust = 0.5),
                         barwidth = 18,
                         label.position = "bottom",
                         label.theme = element_text(size = 14)
                         )) +
  theme(legend.title = element_text(), legend.position = "bottom")

# Add gheatmap 2 for BNF

pgp3_notip_sds <- pgp2_notip_sds + new_scale_fill()

pgp4_notip_sds <- gheatmap(pgp3_notip_sds, smash_sds_cmp[, "BNF", drop=F],
                 offset = 0.215,
                 width=0.05,
                 colnames_angle = 0,
                 colnames_position = "top",
                 colnames_offset_y = 0.9,
                 font.size = 3,
                 color = "grey50") + 
  scale_fill_viridis_c(option="inferno",
                       direction = -1,
                       name = "BNF (15N/14N)",
                       limits = c(0,5),
                       breaks = c(0, 1, 2, 3, 4, 5),
                       guide = guide_colorbar(
                         direction = "horizontal",
                         title.position = "top",
                         title.theme = element_text(size = 18, hjust = 0.5),
                         barwidth = 18,
                         label.position = "bottom",
                         label.theme = element_text(size = 14)
                       )) +
  theme(legend.title = element_text(), legend.position = "bottom")

# Add gheatmap 3 for IAA
pgp5_notip_sds <- pgp4_notip_sds + new_scale_fill()

pgp6_notip_sds <- gheatmap(pgp5_notip_sds, smash_sds_cmp[, "IAA", drop=F],
                    offset = 0.305,
                    width = 0.05,
                    colnames_angle = 0,
                    colnames_position = "top",
                    colnames_offset_y = 0.9,
                    font.size = 3,
                    color = "grey50") + 
  scale_fill_viridis_c(option = "plasma",
                       direction = 1,
                       name = "IAA Biosynthesis (mg/mL)",
                       limits = c(-5,215),
                       breaks = c(0, 50, 100, 150, 200),
                       guide = guide_colorbar(
                         direction = "horizontal",
                         title.position = "top",
                         title.theme = element_text(size = 18, hjust = 0.5),
                         barwidth = 18,
                         label.position = "bottom",
                         label.theme = element_text(size = 14)
                       )) +
  theme(legend.title = element_text(), legend.position = "bottom")

# Add gheatmap 4 for PO4
pgp7_notip_sds <- pgp6_notip_sds + new_scale_fill()

pgp_sds_notip <- gheatmap(pgp7_notip_sds, smash_sds_cmp[, "PO4", drop=F],
                    offset = 0.395,
                    width = 0.05,
                    colnames_angle = 0,
                    colnames_position = "top",
                    colnames_offset_y = 0.9,
                    font.size = 3,
                    color = "grey50") + 
  scale_fill_viridis_c(option = "magma",
                       direction = 1,
                       name = "PO4 Solubilization (mg/L)",
                       limits = c(0,2250),
                       breaks = c(0, 500, 1000, 1500, 2000),
                       guide = guide_colorbar(
                         direction = "horizontal",
                         title.position = "top",
                         title.theme = element_text(size = 18, hjust = 0.5),
                         barwidth = 18,
                         label.position = "bottom",
                         label.theme = element_text(size = 14)
                       )) +
  theme(legend.title = element_text(),
        legend.position = "bottom")

pgp_sds_rot <- print(rotate_tree(pgp_sds_notip, -85))

# save plot
ggsave("./Plots/pgp_sds_rot.pdf", pgp_sds_rot, width = 22, height = 14)

```

##### DSN Isolates
```{r, message=FALSE}
# Add gheatmap 0 for genus
## reproducible colors
pgp0_notip_dsn <- gheatmap(sm_dsn_k31_tree, smash_dsn_cmp[, "Genus", drop = F],
                     offset = 0,
                     width = 0.05,
                     colnames_position = "top",
                     colnames_angle = 0,
                     colnames_offset_y = 4.5,
                     font.size = 2.85,
                     color = "grey50") + 
  scale_fill_manual(values=c("#CAD773",
                             "#81D870",
                             "#E25F9D",
                             "#D8E9D6",
                             "#72A890",
                             "#CDEC4C",
                             "#81748E",
                             "#D3D0E3",
                             "#7DAEDE",
                             "#E2684F",
                             "#E1ABAA",
                             "#7BEFDF",
                             "#D7A7DF",
                             "#E5DEA8",
                             "#CF7DDC"),
                    name = "Genus",
                    guide = guide_legend(
                      direction = "vertical",
                      ncol = 2,
                      title.position = "top",
                      title.theme = element_text(size = 18, hjust = 0.5),
                      label.position = "right",
                      label.theme = element_text(size = 14)
                      )
  ) +
  theme(legend.title = element_text(), legend.key.size = unit(.5, "cm"))

# Add gheatmap 1 for ACC
pgp1_notip_dsn <- pgp0_notip_dsn + new_scale_fill()

pgp2_notip_dsn <- gheatmap(pgp1_notip_dsn, smash_dsn_cmp[, "ACC", drop = F],
                     offset = 0.22,
                     width = 0.05,
                     colnames_position = "top",
                     colnames_angle = 0,
                     colnames_offset_y = 4.5,
                     font.size = 2.85,
                     color = "grey50") + 
  scale_fill_viridis_c(option="viridis",
                       direction = -1,
                       name = "ACC (Relative Growth Rate)",
                       limits = c(0,5),
                       breaks = c(0, 1, 2, 3, 4, 5),
                       guide = guide_colorbar(
                         direction = "horizontal",
                         title.position = "top",
                         title.theme = element_text(size = 18, hjust = 0.5),
                         barwidth = 18,
                         label.position = "bottom",
                         label.theme = element_text(size = 14)
                         )) +
  theme(legend.title = element_text(), legend.position = "bottom")

# Add gheatmap 2 for BNF

pgp3_notip_dsn <- pgp2_notip_dsn + new_scale_fill()

pgp4_notip_dsn <- gheatmap(pgp3_notip_dsn, smash_dsn_cmp[, "BNF", drop=F],
                 offset = 0.37,
                 width=0.05,
                 colnames_angle = 0,
                 colnames_position = "top",
                 colnames_offset_y = 4.5,
                 font.size = 2.85,
                 color = "grey50") + 
  scale_fill_viridis_c(option="inferno",
                       direction = -1,
                       name = "BNF (15N/14N)",
                       limits = c(0,5),
                       breaks = c(0, 1, 2, 3, 4, 5),
                       guide = guide_colorbar(
                         direction = "horizontal",
                         title.position = "top",
                         title.theme = element_text(size = 18, hjust = 0.5),
                         barwidth = 18,
                         label.position = "bottom",
                         label.theme = element_text(size = 14)
                       )) +
  theme(legend.title = element_text(), legend.position = "bottom")

# Add gheatmap 3 for IAA
pgp5_notip_dsn <- pgp4_notip_dsn + new_scale_fill()

pgp6_notip_dsn <- gheatmap(pgp5_notip_dsn, smash_dsn_cmp[, "IAA", drop=F],
                    offset = 0.52,
                    width = 0.05,
                    colnames_angle = 0,
                    colnames_position = "top",
                    colnames_offset_y = 4.5,
                    font.size = 2.85,
                    color = "grey50") + 
  scale_fill_viridis_c(option = "plasma",
                       direction = 1,
                       name = "IAA Biosynthesis (mg/mL)",
                       limits = c(-5,215),
                       breaks = c(0, 50, 100, 150, 200),
                       guide = guide_colorbar(
                         direction = "horizontal",
                         title.position = "top",
                         title.theme = element_text(size = 18, hjust = 0.5),
                         barwidth = 18,
                         label.position = "bottom",
                         label.theme = element_text(size = 14)
                       )) +
  theme(legend.title = element_text(), legend.position = "bottom")

# Add gheatmap 4 for PO4
pgp7_notip_dsn <- pgp6_notip_dsn + new_scale_fill()

pgp_dsn_notip <- gheatmap(pgp7_notip_dsn, smash_dsn_cmp[, "PO4", drop=F],
                    offset = 0.67,
                    width = 0.05,
                    colnames_angle = 0,
                    colnames_position = "top",
                    colnames_offset_y = 4.5,
                    font.size = 2.85,
                    color = "grey50") + 
  scale_fill_viridis_c(option = "magma",
                       direction = 1,
                       name = "PO4 Solubilization (mg/L)",
                       limits = c(0, 2250),
                       breaks = c(0, 500, 1000, 1500, 2000),
                       guide = guide_colorbar(
                         direction = "horizontal",
                         title.position = "top",
                         title.theme = element_text(size = 18, hjust = 0.5),
                         barwidth = 18,
                         label.position = "bottom",
                         label.theme = element_text(size = 14)
                       )) +
  theme(legend.title = element_text(),
        legend.position = "bottom")

pgp_dsn_rot <- print(rotate_tree(pgp_dsn_notip, -83))

# save plot
ggsave("./Plots/pgp_dsn_rot.pdf", pgp_dsn_rot, width = 22, height = 14)

```

### Extract legends for pgp plots
#### Genus
```{r}
# Genus Tree
cowplot_genus <- gheatmap(sm_k31_tree_o15, pgp_master[, "Genus", drop = F],
                     offset = 0,
                     width = 0.05,
                     colnames_position = "top",
                     colnames_angle = 0,
                     colnames_offset_y = 9.5,
                     font.size = 3.5,
                     color = "grey50") + 
  scale_fill_manual(values=c("#E6BE53",
                             "#CAD773",
                             "#81D870",
                             "#8F3EE4",
                             "#E25F9D",
                             "#D8E9D6",
                             "#72A890",
                             "#CDEC4C",
                             "#C59D6B",
                             "#86D6E1",
                             "#77E94F",
                             "#81748E",
                             "#64E7A9",
                             "#D3D0E3",
                             "#767CDE",
                             "#7DAEDE",
                             "#E2684F",
                             "#E1ABAA",
                             "#7BEFDF",
                             "#DA47D2",
                             "#D7A7DF",
                             "#E5DEA8",
                             "#B4E3A7",
                             "#CF7DDC"),
                    name = "Genus",
                    guide = guide_legend(
                      direction = "vertical",
                      ncol = 3,
                      title.position = "top",
                      title.theme = element_text(size = 18, hjust = 0.5),
                      label.position = "right",
                      label.theme = element_text(size = 16)
                      )) + theme(legend.title = element_text(),
                                 legend.key.size = unit(0.75, "cm"))

cowplot_genus

legend_genus <- get_legend(cowplot_genus)
```

#### ACC
```{r}
cowplot_acc <- gheatmap(sm_dsp_k31_tree, pgp_master[, "ACC", drop = F],
                     offset = 00,
                     width = 0.05,
                     colnames_position = "top",
                     colnames_angle = 0,
                     colnames_offset_y = 2.25,
                     font.size = 3.5,
                     color = "grey50") + 
  scale_fill_viridis_c(option="viridis",
                       direction = -1,
                       name = "ACC (Relative Growth Rate)",
                       limits = c(0,5),
                       breaks = c(0, 1, 2, 3, 4, 5),
                       guide = guide_colorbar(
                         direction = "horizontal",
                         title.position = "top",
                         title.theme = element_text(size = 18, hjust = 0.5),
                         barwidth = 18,
                         label.position = "bottom",
                         label.theme = element_text(size = 14)
                         )) +
  theme(legend.title = element_text(), legend.position = "bottom")

cowplot_acc

legend_acc <- get_legend(cowplot_acc)

```

#### BNF
```{r}
cowplot_bnf <- gheatmap(sm_dsp_k31_tree, pgp_master[, "BNF", drop = F],
                     offset = 00,
                     width = 0.05,
                     colnames_position = "top",
                     colnames_angle = 0,
                     colnames_offset_y = 2.25,
                     font.size = 3.5,
                     color = "grey50") + 
  scale_fill_viridis_c(option="inferno",
                       direction = -1,
                       name = "BNF (15N/14N Ratio)",
                       limits = c(0,5),
                       breaks = c(0, 1, 2, 3, 4, 5),
                       guide = guide_colorbar(
                         direction = "horizontal",
                         title.position = "top",
                         title.theme = element_text(size = 18, hjust = 0.5),
                         barwidth = 18,
                         label.position = "bottom",
                         label.theme = element_text(size = 14)
                         )) +
  theme(legend.title = element_text(), legend.position = "bottom")

cowplot_bnf

legend_bnf <- get_legend(cowplot_bnf)

```

#### IAA
```{r}
cowplot_iaa <- gheatmap(sm_dsp_k31_tree, pgp_master[, "IAA", drop = F],
                     offset = 0,
                     width = 0.05,
                     colnames_position = "top",
                     colnames_angle = 0,
                     colnames_offset_y = 2.25,
                     font.size = 3.5,
                     color = "grey50") + 
  scale_fill_viridis_c(option = "plasma",
                       direction = 1,
                       name = "IAA (mg/mL)",
                       limits = c(-5, 215),
                       breaks = c(0, 50, 100, 150, 200),
                       guide = guide_colorbar(
                         direction = "horizontal",
                         title.position = "top",
                         title.theme = element_text(size = 18, hjust = 0.5),
                         barwidth = 18,
                         label.position = "bottom",
                         label.theme = element_text(size = 14)
                         )
                       ) +
  theme(legend.title = element_text(), legend.position = "bottom")

cowplot_iaa

legend_iaa <- get_legend(cowplot_iaa)

```

#### PO4
```{r}
cowplot_po4 <- gheatmap(sm_dsp_k31_tree, pgp_master[, "PO4", drop = F],
                     offset = 0,
                     width = 0.05,
                     colnames_position = "top",
                     colnames_angle = 0,
                     colnames_offset_y = 2.25,
                     font.size = 3.5,
                     color = "grey50") + 
  scale_fill_viridis_c(option = "magma",
                       direction = 1,
                       name = "PO4 (mg/L)",
                       limits = c(0, 2250),
                       breaks = c(0, 500, 1000, 1500, 2000),
                       guide = guide_colorbar(
                         direction = "horizontal",
                         title.position = "top",
                         title.theme = element_text(size = 18, hjust = 0.5),
                         barwidth = 18,
                         label.position = "bottom",
                         label.theme = element_text(size = 14)
                         )
                       ) +
  theme(legend.title = element_text(), legend.position = "bottom")

cowplot_po4

legend_po4 <- get_legend(cowplot_po4)

```


### Assay Legend Array
```{r}

pgp_legend_array_1 <- plot_grid(legend_acc,
                              legend_bnf,
                              nrow = 2)

pgp_legend_array_2 <- plot_grid(legend_iaa,
                                legend_po4,
                                nrow = 2)

pgp_legend_array <- plot_grid(pgp_legend_array_1,
                              pgp_legend_array_2,
                              ncol = 2)

pgp_all_legend <- plot_grid(legend_genus,
                            pgp_legend_array,
                            nrow = 2)
pgp_all_legend

```

#### Plot Grid


```{r}
pgp_row <- plot_grid(pgp_dsp_rot + theme(plot.margin = unit(c(0, 0, 0, 0), "cm"),
                                         legend.position = "none"),
                     pgp_sds_rot + theme(plot.margin = unit(c(0, 0, 0, 0), "cm"),
                                         legend.position = "none"),
                     pgp_dsn_rot + theme(plot.margin = unit(c(0, 0, 0, 0), "cm"),
                                         legend.position = "none"),
                     pgp_all_legend,
                     labels = c("A", "B", "C", ""),
                     label_size = 24,
                     nrow = 2,
                     ncol = 2,
                     vjust = 3,
                     align = 'vh'
                     )

#pgp_grid <- plot_grid(pgp_row, pgp_all_legend,
#                    ncol = 1,
#                    rel_heights = c(1, 0.5))
pgp_row

ggsave("./plots/Fig1.pdf", pgp_row, width = 22, height = 14)
```



## Min and Max Phenotype values
```{r}
print("BNF")
min(pgp_master[,"BNF"], na.rm = TRUE)
max(pgp_master[,"BNF"], na.rm = TRUE)

print("ACC")
min(pgp_master[,"ACC"], na.rm = TRUE)
max(pgp_master[,"ACC"], na.rm = TRUE)

print("IAA")
min(pgp_master[,"IAA"], na.rm = TRUE)
max(pgp_master[,"IAA"], na.rm = TRUE)

print("PO4")
min(pgp_master[,"PO4"], na.rm = TRUE)
max(pgp_master[,"PO4"], na.rm = TRUE)
```


## Data Summary

### BNF Assay
#### DSP Group
```{r}
# DSP Group
## BNF < 1
smash_dsp_cmp %>%
  group_by(Genus, BNF) %>%
  select(BNF) %>%
  filter(BNF < 1) %>%
  count()

## 1 <= BNF <= 2
smash_dsp_cmp %>%
  group_by(Genus, BNF) %>%
  select(BNF) %>%
  filter(BNF >= 1 & BNF <=2) %>%
  count()

## BNF >= 2
smash_dsp_cmp %>%
  group_by(Genus, BNF) %>%
  select(BNF) %>%
  filter(BNF >= 2) %>%
  count()

stable_dsp_bnf_rnk <-
smash_dsp_cmp %>%
  group_by(Genus, BNF) %>%
  arrange(desc(BNF)) %>%
  select("bcw_id", "Genus", "BNF") %>%
  mutate(NIF_Group = "DSP")
```

#### SDS Group
```{r}
# SDS Group
## BNF < 1
smash_sds_cmp %>%
  group_by(Genus, BNF) %>%
  select(BNF) %>%
  filter(BNF < 1) %>%
  count()

## 1 <= BNF <= 2
smash_sds_cmp %>%
  group_by(Genus, BNF) %>%
  select(BNF) %>%
  filter(BNF >= 1 & BNF <2) %>%
  count()

## BNF >= 2
smash_sds_cmp %>%
  group_by(Genus, BNF) %>%
  select(BNF) %>%
  filter(BNF >= 2) %>%
  count()

stable_sds_bnf_rnk <-
smash_sds_cmp %>%
  group_by(Genus, BNF) %>%
  arrange(desc(BNF)) %>%
  select("bcw_id", "Genus", "BNF") %>%
  mutate(NIF_Group = "SDS")
```

#### DSN Group
```{r}
# DSN Group
## BNF < 1
smash_dsn_cmp %>%
  group_by(Genus, BNF) %>%
  select(BNF) %>%
  filter(BNF < 1) %>%
  count()

## 1 <= BNF <= 2
smash_dsn_cmp %>%
  group_by(Genus, BNF) %>%
  select(BNF) %>%
  filter(BNF >= 1 & BNF <2) %>%
  count()

## BNF >= 2
smash_dsn_cmp %>%
  group_by(Genus, BNF) %>%
  select(BNF) %>%
  filter(BNF >= 2) %>%
  count()

stable_dsn_bnf_rnk <-
smash_dsn_cmp %>%
  group_by(Genus, BNF) %>%
  arrange(desc(BNF)) %>%
  select("bcw_id", "Genus", "BNF") %>%
  mutate(NIF_Group = "DSN")
```

### ACC Assay
#### DSP Group
```{r}
# DSP Group
## ACC < 1
smash_dsp_cmp %>%
  group_by(Genus, ACC) %>%
  select(ACC) %>%
  filter(ACC < 1) %>%
  count()

## 1 <= ACC <= 2
smash_dsp_cmp %>%
  group_by(Genus, ACC) %>%
  select(ACC) %>%
  filter(ACC >= 1 & ACC <=2) %>%
  count()

## 2 <= ACC <= 3
smash_dsp_cmp %>%
  group_by(Genus, ACC) %>%
  select(ACC) %>%
  filter(ACC >= 2 & ACC <= 3) %>%
  count()

## 3 <= ACC <= 4
smash_dsp_cmp %>%
  group_by(Genus, ACC) %>%
  select(ACC) %>%
  filter(ACC >= 3 & ACC <= 4) %>%
  count()

## ACC < 4
smash_dsp_cmp %>%
  group_by(Genus, ACC) %>%
  select(ACC) %>%
  filter(ACC > 4) %>%
  count()

## ACC RANKED
colnames(smash_dsp_cmp[,194:200])

stable_dsp_acc_rnk <-
smash_dsp_cmp %>%
  group_by(Genus, ACC) %>%
  arrange(desc(ACC)) %>%
  select("bcw_id", "Genus", "ACC") %>%
  mutate(NIF_Group = "DSP")
```

#### SDS Group
```{r}
# SDS Group
## ACC < 1
smash_sds_cmp %>%
  group_by(Genus, ACC) %>%
  select(ACC) %>%
  filter(ACC < 1) %>%
  count()

## 1 <= ACC <= 2
smash_sds_cmp %>%
  group_by(Genus, ACC) %>%
  select(ACC) %>%
  filter(ACC >= 1 & ACC <=2) %>%
  count()

## 2 <= ACC <= 3
smash_sds_cmp %>%
  group_by(Genus, ACC) %>%
  select(ACC) %>%
  filter(ACC >= 2 & ACC <= 3) %>%
  count()

## 3 <= ACC <= 4
smash_sds_cmp %>%
  group_by(Genus, ACC) %>%
  select(ACC) %>%
  filter(ACC >= 3 & ACC <= 4) %>%
  count()

## ACC < 4
smash_sds_cmp %>%
  group_by(Genus, ACC) %>%
  select(ACC) %>%
  filter(ACC > 4) %>%
  count()

## ACC RANKED
colnames(smash_sds_cmp[,67:73])

stable_sds_acc_rnk <-
smash_sds_cmp %>%
  group_by(Genus, ACC) %>%
  arrange(desc(ACC)) %>%
  select("bcw_id", "Genus", "ACC") %>%
  mutate(NIF_Group = "SDS")
```

#### DSN Group
```{r}
# DSN Group
## ACC < 1
smash_dsn_cmp %>%
  group_by(Genus, ACC) %>%
  select(ACC) %>%
  filter(ACC < 1) %>%
  count()

## 1 <= ACC <= 2
smash_dsn_cmp %>%
  group_by(Genus, ACC) %>%
  select(ACC) %>%
  filter(ACC >= 1 & ACC <=2) %>%
  count()

## 2 <= ACC <= 3
smash_dsn_cmp %>%
  group_by(Genus, ACC) %>%
  select(ACC) %>%
  filter(ACC >= 2 & ACC <= 3) %>%
  count()

## 3 <= ACC <= 4
smash_dsn_cmp %>%
  group_by(Genus, ACC) %>%
  select(ACC) %>%
  filter(ACC >= 3 & ACC <= 4) %>%
  count()

## ACC < 4
smash_dsn_cmp %>%
  group_by(Genus, ACC) %>%
  select(ACC) %>%
  filter(ACC > 4) %>%
  count()

## ACC RANKED
colnames(smash_dsn_cmp[,234:240])

stable_dsn_acc_rnk <-
smash_dsn_cmp %>%
  group_by(Genus, ACC) %>%
  arrange(desc(ACC)) %>%
  select("bcw_id", "Genus", "ACC") %>%
  mutate(NIF_Group = "DSN")
```


### IAA Assay
#### DSP Group
```{r}
# DSP Group
## IAA < 1
smash_dsp_cmp %>%
  group_by(Genus, IAA) %>%
  select(IAA) %>%
  filter(IAA < 1) %>%
  count()

## 1 <= IAA <= 2
smash_dsp_cmp %>%
  group_by(Genus, IAA) %>%
  select(IAA) %>%
  filter(IAA >= 1 & IAA <=2) %>%
  count()

## 2 <= IAA <= 3
smash_dsp_cmp %>%
  group_by(Genus, IAA) %>%
  select(IAA) %>%
  filter(IAA >= 2 & IAA <= 3) %>%
  count()

## 3 <= IAA <= 4
smash_dsp_cmp %>%
  group_by(Genus, IAA) %>%
  select(IAA) %>%
  filter(IAA >= 3 & IAA <= 4) %>%
  count()

## IAA < 4
smash_dsp_cmp %>%
  group_by(Genus, IAA) %>%
  select(IAA) %>%
  filter(IAA > 4) %>%
  count()

## IAA RANKED
colnames(smash_dsp_cmp[,194:200])

stable_dsp_iaa_rnk <-
smash_dsp_cmp %>%
  group_by(Genus, IAA) %>%
  arrange(desc(IAA)) %>%
  select("bcw_id", "Genus", "IAA") %>%
  mutate(NIF_Group = "DSP")
```

#### SDS Group
```{r}
# SDS Group
## IAA < 1
smash_sds_cmp %>%
  group_by(Genus, IAA) %>%
  select(IAA) %>%
  filter(IAA < 1) %>%
  count()

## 1 <= IAA <= 2
smash_sds_cmp %>%
  group_by(Genus, IAA) %>%
  select(IAA) %>%
  filter(IAA >= 1 & IAA <=2) %>%
  count()

## 2 <= IAA <= 3
smash_sds_cmp %>%
  group_by(Genus, IAA) %>%
  select(IAA) %>%
  filter(IAA >= 2 & IAA <= 3) %>%
  count()

## 3 <= IAA <= 4
smash_sds_cmp %>%
  group_by(Genus, IAA) %>%
  select(IAA) %>%
  filter(IAA >= 3 & IAA <= 4) %>%
  count()

## IAA < 4
smash_sds_cmp %>%
  group_by(Genus, IAA) %>%
  select(IAA) %>%
  filter(IAA > 4) %>%
  count()

## IAA RANKED
colnames(smash_sds_cmp[,67:73])

stable_sds_iaa_rnk <- 
smash_sds_cmp %>%
  group_by(Genus, IAA) %>%
  arrange(desc(IAA)) %>%
  select("bcw_id", "Genus", "IAA") %>%
  mutate(NIF_Group = "SDS")
```

#### DSN Group
```{r}
# DSN Group
## IAA < 1
smash_dsn_cmp %>%
  group_by(Genus, IAA) %>%
  select(IAA) %>%
  filter(IAA < 1) %>%
  count()

## 1 <= IAA <= 2
smash_dsn_cmp %>%
  group_by(Genus, IAA) %>%
  select(IAA) %>%
  filter(IAA >= 1 & IAA <=2) %>%
  count()

## 2 <= IAA <= 3
smash_dsn_cmp %>%
  group_by(Genus, IAA) %>%
  select(IAA) %>%
  filter(IAA >= 2 & IAA <= 3) %>%
  count()

## 3 <= IAA <= 4
smash_dsn_cmp %>%
  group_by(Genus, IAA) %>%
  select(IAA) %>%
  filter(IAA >= 3 & IAA <= 4) %>%
  count()

## IAA < 4
smash_dsn_cmp %>%
  group_by(Genus, IAA) %>%
  select(IAA) %>%
  filter(IAA > 4) %>%
  count()

## IAA RANKED
colnames(smash_dsn_cmp[,234:240])

stable_dsn_iaa_rnk <-
smash_dsn_cmp %>%
  group_by(Genus, IAA) %>%
  arrange(desc(IAA)) %>%
  select("bcw_id", "Genus", "IAA") %>%
  mutate(NIF_Group = "DSN")
```

### PO4 Assay
#### DSP Group
```{r}
# DSP Group
## PO4 < 1
smash_dsp_cmp %>%
  group_by(Genus, PO4) %>%
  select(PO4) %>%
  filter(PO4 < 1) %>%
  count()

## 1 <= PO4 <= 2
smash_dsp_cmp %>%
  group_by(Genus, PO4) %>%
  select(PO4) %>%
  filter(PO4 >= 1 & PO4 <=2) %>%
  count()

## 2 <= PO4 <= 3
smash_dsp_cmp %>%
  group_by(Genus, PO4) %>%
  select(PO4) %>%
  filter(PO4 >= 2 & PO4 <= 3) %>%
  count()

## 3 <= PO4 <= 4
smash_dsp_cmp %>%
  group_by(Genus, PO4) %>%
  select(PO4) %>%
  filter(PO4 >= 3 & PO4 <= 4) %>%
  count()

## PO4 < 4
smash_dsp_cmp %>%
  group_by(Genus, PO4) %>%
  select(PO4) %>%
  filter(PO4 > 4) %>%
  count()

## PO4 RANKED
colnames(smash_dsp_cmp[,194:200])

stable_dsp_po4_rnk <-
smash_dsp_cmp %>%
  group_by(Genus, PO4) %>%
  arrange(desc(PO4)) %>%
  select("bcw_id", "Genus", "PO4") %>%
  mutate(NIF_Group = "DSP")
```

#### SDS Group
```{r}
# SDS Group
## PO4 < 1
smash_sds_cmp %>%
  group_by(Genus, PO4) %>%
  select(PO4) %>%
  filter(PO4 < 1) %>%
  count()

## 1 <= PO4 <= 2
smash_sds_cmp %>%
  group_by(Genus, PO4) %>%
  select(PO4) %>%
  filter(PO4 >= 1 & PO4 <=2) %>%
  count()

## 2 <= PO4 <= 3
smash_sds_cmp %>%
  group_by(Genus, PO4) %>%
  select(PO4) %>%
  filter(PO4 >= 2 & PO4 <= 3) %>%
  count()

## 3 <= PO4 <= 4
smash_sds_cmp %>%
  group_by(Genus, PO4) %>%
  select(PO4) %>%
  filter(PO4 >= 3 & PO4 <= 4) %>%
  count()

## PO4 < 4
smash_sds_cmp %>%
  group_by(Genus, PO4) %>%
  select(PO4) %>%
  filter(PO4 > 4) %>%
  count()

## PO4 RANKED
colnames(smash_sds_cmp[,67:73])

stable_sds_po4_rnk <-
smash_sds_cmp %>%
  group_by(Genus, PO4) %>%
  arrange(desc(PO4)) %>%
  select("bcw_id", "Genus", "PO4") %>%
  mutate(NIF_Group = "SDS")
```

#### DSN Group
```{r}
# DSN Group
## PO4 < 1
smash_dsn_cmp %>%
  group_by(Genus, PO4) %>%
  select(PO4) %>%
  filter(PO4 < 1) %>%
  count()

## 1 <= PO4 <= 2
smash_dsn_cmp %>%
  group_by(Genus, PO4) %>%
  select(PO4) %>%
  filter(PO4 >= 1 & PO4 <=2) %>%
  count()

## 2 <= PO4 <= 3
smash_dsn_cmp %>%
  group_by(Genus, PO4) %>%
  select(PO4) %>%
  filter(PO4 >= 2 & PO4 <= 3) %>%
  count()

## 3 <= PO4 <= 4
smash_dsn_cmp %>%
  group_by(Genus, PO4) %>%
  select(PO4) %>%
  filter(PO4 >= 3 & PO4 <= 4) %>%
  count()

## PO4 < 4
smash_dsn_cmp %>%
  group_by(Genus, PO4) %>%
  select(PO4) %>%
  filter(PO4 > 4) %>%
  count()

## PO4 RANKED
colnames(smash_dsn_cmp[,234:240])

stable_dsn_po4_rnk <-
smash_dsn_cmp %>%
  group_by(Genus, PO4) %>%
  arrange(desc(PO4)) %>%
  select("bcw_id", "Genus", "PO4") %>%
  mutate(NIF_Group = "DSN")
```

#### Save Sup. Tables
```{r}
# create list of dataframes
sup_list <- mget(ls(pattern = "stable_.+"))

sup_path_out <- "sup_tables/"

for(i in names(sup_list)){
  write.csv(sup_list[[i]], paste(sup_path_out, i,".csv"))
}
```
